---
title: "Pleural protein dilution series"
subtitle: "Results from Olink Target 96 Inflammation panel"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = F, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("dilution-series-qc.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Background 

16 samples from previous pleural fluid olink inflammation analysis (""GB390725-RB_pleural fluid_NPX.csv") was re-run at several different concentration dilutions in order to find the optimal conditions for further analysis.

## Design

Counts by dilution fraction below:

```{r globals, results = 'hide', echo=F }
```

```{r load.olink}
```

```{r load.elisa}
```

```{r access.pheno}
```
```{r dilutions}
```

`patient.ids` in each dilution:

```{r dilution.ids, echo = T}
```
# Frequency below LOD by dilution condition

The bar chart below shows, for each protein, the number of measurements falling below the limit of detection (LOD) at each dilution level. Proteins are ordered by their total count of below-LOD values. A high frequency of below-LOD measurements at a given dilution suggests that the sample concentration may be too low for reliable quantification of that protein.

```{r lod, fig.width = 10, fig.height = 20}
```

# Correlation across dilution conditions for all proteins

To assess measurement consistency across dilution levels, Pearson correlations were computed for each protein between every pair of dilution conditions. For each pair, only individuals with measurements at both dilution levels were included. 

The table below summarises the median and mean per-protein correlation for each dilution pair, along with the number of individuals and proteins contributing to each comparison.

```{r dilution-pair-correlations, results = 'asis'}
```

The violin and box plot below shows the full distribution of per-protein correlations for each dilution pair.

```{r dilution-pair-correlations-plot, fig.width = 8, fig.height = 5}
```

# Pairwise dilution scatter plots for selected proteins

To investigate how individual protein measurements track across dilution levels, scatter plots are shown below for a panel of ten biologically relevant proteins: IL-6, TNF, IL-8, uPA, CXCL1, IFN-γ, CXCL5, IL-17A, IL-1α, and CCL20. These proteins were selected a priori as the key protein targets to have in assay range.

For each protein, every pairwise combination of dilution levels is displayed as a separate panel. Each point represents one individual, with the x- and y-axes showing the NPX value at the two dilution levels being compared. The blue line shows the ordinary least-squares linear fit; the dashed grey line is the identity line (y = x), indicating perfect agreement. The Pearson correlation coefficient (r) is annotated in the top-left corner of each panel.

```{r dilution-scatter-fn, echo = FALSE}
```

```{r dilution-scatter-plots, fig.width = 12, fig.height = 16, results = 'asis'}
```

