---
title: "Pleural protein dilution series"
subtitle: "Results from Olink Target 96 Inflammation panel"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = F, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("dilution-series-qc.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Background 

16 samples from previous pleural fluid olink inflammation analysis (""GB390725-RB_pleural fluid_NPX.csv") was re-run at several different concentration dilutions in order to find the optimal conditions for further analysis.

## Design

Counts by dilution fraction below:

```{r globals, results = 'hide', echo=F }
```

```{r load.data}
```

```{r access.pheno}
```
```{r dilutions}
```

`patient.ids` in each dilution:

```{r dilution.ids, echo = T}
```

# Correlation across dilution conditions for all proteins

To assess measurement consistency across dilution levels, Pearson correlations were computed for each protein between every pair of dilution conditions. For each pair, only individuals with measurements at both dilution levels were included. 

The table below summarises the median and mean per-protein correlation for each dilution pair, along with the number of individuals and proteins contributing to each comparison.

```{r dilution-pair-correlations, results = 'asis'}
```

The violin and box plot below shows the full distribution of per-protein correlations for each dilution pair.

```{r dilution-pair-correlations-plot, fig.width = 8, fig.height = 5}
```

