---
title: "Comparison of IL6 measures in study of pleural infection"
subtitle: "Results from Olink Target 96 Inflammation panel "
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = T, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("il6-compare.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Samples measured by ELISA and Olink

## Fergus Hamilton's serum and pleural ELISA IL6 measurements
- **il6_clean_los.txt**: ELISA Measured pleural IL6 as received from Fergus Hamilton by email on 24 Nov 2025


```{r globals, results = 'hide', echo=F }
```

```{r load.elisa, echo=T}
```

```{r check.elisa, results = 'hide', echo=F}
```

## Olink Pleural and Plasma Target 96 measurements

3 plates of pleural Olink Target 96 assays were run in Nov 2025
* full qc details here: [pleural_metaboprep_qc_report.html](https://yousefi138.github.io/pleural-proteomics/pleural_metaboprep_qc_report.html)

At the same time, 1 plate of plasma Olink Target 96 assays were run. These samples were all overlapping with those selected for pleural analysis.
* full qc details here: [plasma_metaboprep_qc_report.html](https://yousefi138.github.io/pleural-proteomics/plasma_metaboprep_qc_report.html)


```{r load.olink, results = 'hide', echo=F}
```

```{r check.olink, results = 'hide', echo=F}
```

Overlap of olink samples with elisa samples: 

```{r overlap}
```

## Measurement counts

```{r join}
```

```{r count}
```

# Distributions

```{r make.long, results = 'hide', echo=F}
```
```{r dist, results = 'hide', echo=F}
```

```{r fig.dist, fig.width = 10, fig.height = 10}
```






