---
title: "Pleural protein 16x dilution plate 01"
subtitle: "Results from Olink Target 96 Inflammation panel"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = F, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("qc-16x-dilution01.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Background 

Some proteins in our previous pleural fluid Olink Target 96 Inflammation pannel analysis (""GB390725-RB_pleural fluid_NPX.csv") were observed to be outside the dynamic range of this assay. In particular, we were most confident in this behaviour for the IL6 protein abundance data, as we had independent high quality ELISA data from these same samples which showed poor correlation with the Olink measures. We think that this poor correlation was driven by the IL6 levels in our pleural samples being greatly above the dynamic range of the Olink assay.

To assess whether our pleural protein abundances are within the dynamic range of the olink assay, we have re-run a subset of 16 samples from previous pleural fluid olink inflammation analysis (""GB390725-RB_pleural fluid_NPX.csv") at several different concentration dilutions.

Based on these results summarised in [dilution-series-qc.html](https://yousefi138.github.io/pleural-proteomics/dilution-series-qc.html), we selected a 16x dilution and ran 1 plate of samples ("GB390725-RB_1_16_Extended_2026-07-06.csv") before running the assay on the entire sample set at 16x dilution. 

## Purpose


This project to verify whether 16x is still an appropriate/optimal dilution concentration for future olink assays of pleural fluid. As it may not be possible for all Olink Target 96 Inflammation pannel protein targets to be in range, we have a set of 10 key highest priority protein targets where we would like the assay to demonstrate good performance:

> IL-6, TNF, IL-8, uPA, CXCL1, IFN-γ, CXCL5, IL-17A, IL-1α, and CCL20

IL6 is our highest priority target as well as the one protein where we have independent gold standard ELISA data for comparison in matched pleural and serum samples. 

```{r globals, results = 'hide', echo=F }
```

```{r load.olink}
```

```{r load.elisa}
```

```{r il6}
```

# QC of all measured proteins

## Frequency below LOD by dilution condition

The bar chart below shows, for each protein, the number of measurements falling below the limit of detection (LOD) at each dilution level. Proteins are ordered by their total count of below-LOD values. A high frequency of below-LOD measurements at a given dilution suggests that the sample concentration may be too low for reliable quantification of that protein.

### All proteins

```{r lod, fig.width = 10, fig.height = 20}
```

### High priority proteins

Subset to our 10 high priority proteins

```{r lod.sub, fig.width = 10, fig.height = 20}
```

## Observed NPX distribution by assay/protein and dilution

### High priority proteins

Distributions of NPX for the same protein subset, grouped by assay and dilution.

```{r dist-subset, fig.width = 15, fig.height = 20}
```

### All remaining proteins

Distributions of NPX by dilution for all remaining assays not in the priority set.

```{r dist-other, fig.width = 15, fig.height = 80}
```

# Correlation with ELISA IL6

Number of samples with paired ELISA IL6 measurements:

```{r count}
```

## Linear trendlines

```{r lm, echo=F, fig.width = 10, fig.height = 10}
```



