---
title: "`r study`"
date: "`r format(Sys.time(), '%d %B %Y')`"
author: "`r author`"
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

# Parameters

```{r, results="markup"}
object$parameters
```

# QQ plot

```{r,results="asis"}
object$qq.plot
```

# Volcano plot

```{r,results="asis"}
object$volcano.plot
```

# Significant proteins

Below are the `r length(object$significant.sites)`
proteins with association p-values below the `$sig.threshold` or p-values < `r object$parameters$sig.threshold`.

```{r}
tab <- with(object, mol.stats[match(significant.sites, rownames(mol.stats)),])
```

```{r}
knitr::kable(tab)
```

```{r}
practical.sites <- object$practical.sites
tab <- with(object, mol.stats[match(practical.sites, rownames(mol.stats)),])
```


Below are the `r length(practical.sites)`
proteins with association p-values below the `$practical.threshold` or p-values < `r object$parameters$practical.threshold`
in the `r object$parameters$model` regression model.

```{r}
knitr::kable(tab)
```


# R session information

```{r, results="markup"}
sessionInfo()
```