---
title: "Compare pleural and plasma proteomics of pleural infection"
subtitle: "Results from Olink Target 96 Inflammation panel "
date: "`r format(Sys.time(), '%d %B %Y')`"
author: Paul Yousefi
output:
  html_document:
    css: docs/style.css
    toc: true
    highlight: tango
---

```{r setoptions, eval = TRUE, echo = FALSE}
knitr::opts_chunk$set(
comment = ">",
error = FALSE,
tidy = FALSE,
echo = T, 
warning = FALSE, 
message = FALSE,
cache=FALSE)
```

```{r call_source, echo = F}
path <- "."
sourcecode <- grep("tissue-compare.r$", 
	list.files(path), value = T)
read_chunk(file.path(path, sourcecode))
```

# Prep protein measurements

```{r globals, results = 'hide', echo=F }
```

```{r load.data, echo=T}
```

```{r overlap}
```

Restrict to the N = `r length(intersect(colnames(pleural.all), colnames(plasma.all)))` participants with paired plasma and pleural protein measurements.

```{r restrict}
```

# Compare plasma and pleural protein measurements

## Correlations for all proteins

Full pairwise correlation matrix from all proteins measured in plasma (rows) and pleural fluid (columns):

```{r correlation}
```

```{r cor.mat, fig.width = 14, fig.height = 14}
```

## Correlation distribution

Distribution of the correlations of matched proteins measure in plasma vs. pleural fluid

```{r hist}
```

## Scatter plot 

```{r scatter, echo=F}
```
```{r scatter.fig, echo=F, fig.width = 16, fig.height = 64}
```


